Parallel multi-objective optimization using selective operator application islands

### 2017-07-04 hpmoon-coor.tex
Adding @pacastillo comments and renaming sections according to @JJ. I will submit it this afternoon

In this paper, a new method to
automatically adapt the size of the overlapping, or shared sections,
of the individuals has been compared to other collaborative
evolutionary techniques. Different number of islands, chromosome sizes
and problems have been used. The analysis of the obtained experimental
results, by using different metrics, shows that our approach can provide statistically significant improvements with
respect to the base algorithm.%, being the method to automatically adapt
%the overlapping size the one to obtain the best performance during the
%same execution time. 
 Results also show that the relation between the number
of islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables. 
\end{abstract}

@@

The rest of the paper is organized as follows: after the State of the Art in distributed and co-evolutionary MOEAs, 
the used methodology and the compared algorithms are described in Section \ref{sec:co-evo}. 
Then, the results of the experiments are presented (Section \ref{sec:res}), Finally, the conclusions and future work are discussed.

@@

Since the early 2000s distributed and parallel Evolutionary
Algorithms have been used mostly to leverage systems such
as clusters or grids \citep{Talbi08Parallel}, since a first-order
approach to EA parallelization is as easy as dividing the population
and add a simple communication mechanism among them. But in the case of
MOEAs, the distribution and parallelization is not as direct as in
single-objective EAs. This is because in different steps of the
algorithm a whole set of dependent solutions, the Pareto Front, should
be managed as a whole, spending time in gathering all individuals from
the different processors or islands.

To solve this issue, some authors have proposed using Master-Slave approaches. For example,  \citep{Durillo08masterslave} compared different master-slave approaches: synchronous generational, asynchronous generational and asynchronous steady-state, being the latter the most promising option. The method proposed by Hiroyasu \citep{Hiroyasu07discussion} generates offspring depending on the computational power.

%This is an improvement over the previous? What is the relationship? FERGU: linked
Other kind of approaches have also been explored. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. Their authors concluded that dividing the search space is a good idea, although achieving this is not trivial. The division of the search space has been explored by other researchers, for example, dividing the population in elite and search sub-populations \citep{Wang09parallel}, or separation in processors by objective \citep{Xiao03specialized}. Other authors, such as Martens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

@@

\section{Methodology} % This should be called
                                % methodology. The object of the
                                % section is to explain the
                                % methodology used to test the
                                % problem, in order to do that we have
                                % tested different versions that prove
                                % that our methodology works for the
                                % precise reasons we think it works- JJ FERGU: Done, and explained
\label{sec:co-evo}

@@

In this paper we analyse several SLA algorithms we have
implemented by using a parallel multi-objective baseline algorithm
as a basis for comparisons. Our proposed methods, using different overlapping schemes, %methods or method?
                                %Which methods? - JJ Fergu: clarifying
are based on NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. % This is SoA stuff - JJ FERGU: Yes, I added it here againg because a reviewer wasn't very attentive
Therefore, we have used a basic distributed NSGA-II algorithm without
overlapping sections as the baseline (B).

%This probably should deserve a "Algorithm" environment too - JJ FERGU: You mean the LaTeX one? I've used it in the pseudocode figures
This basic distributed algorithm spreads the population among $P$
islands; after a fixed number of generations,
one individual in a given island is migrated to another random island,
thus avoiding to synchronize the global Pareto Front (PF) every
certain number of generations as other methods described in Section 2
do. At the end of the run, the Pareto Fronts of all islands are
aggregated into a new one, and the quality measures are evaluated. 
Different  SLA alternatives can be devised to evolve the
subpopulations according to the decision space to be explored by each
islands. Here we will consider three methods in which each island only
performs crossover and mutation in specific sections of the
individuals.

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier. Note that, differently from other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel} all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with decomposable and no decomposable problems. %Don't understand what you say
                                %here. What you are saying is that our
                                %approach can't deal with decomposable
                                %problems, actually. - JJ FERGU: you are right, fixed

@@

\subsection{SLA with disjoint islands (D)} 
In this approach, each individual of size $L$ is split into $P$ chunks
of size $L/P$. Every island $p$ only performs crossover and mutation
on the $p_{th}$ part of the individuals. Figure \ref{fig:disjoint}
describes this approach.   
%
\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandDisjoint.jpg}
\caption{SLA with disjoint islands (D): every island $p$ only modifies the $p_{th}$ components (in grey) of the individuals.}
\label{fig:disjoint}
\end{figure*}

\subsection{SLA with overlapping islands (O)}
This approach is similar to the previous one, but every island also uses the $p+1$ and $p-1$ (module size) chunks of the individual for crossover and migration. Therefore, some kind of overlapping of the crossed and mutated parts exists between islands. Figure \ref{fig:overlapping} shows the affected parts of the individuals in
each island.

@@

\subsection{Adaptive SLA with overlapping islands (A)} 
As in the previous method, the sections to address are overlapped, but
instead using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.
% That's not adaptive. It's simply variable - JJ FERGU: it depends on the number of islands. What term do you suggest?
As a first
approach to automatically calculate this value, the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands were $>$8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to create the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. As mentioned, this is only a first approach to obtain $c$, and future work will study new methods to calculate it. Figure \ref{fig:adaptive} explains this method.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%  EXPERIMENTS AND RESULTS  %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\section{Experiments and Results}
\label{sec:res}

This section describes the quality indicators used and the experimental setup. The chosen quality indicators, are:

@@

The ECJ framework \citep{ECJ} has been used to run the
experiments. Specific operators have been developed as new modules for
ECJ, and they can be downloaded from our GitHub repository under a
LGPL V3 License
\footnote{\url{https://github.com/hpmoon/hpmoon-islands}}. 
%\footnote{\url{https://ANONYMOUS-REPOSITORY}}. 
The
island model has been executed asynchronously, using the ECJ distributed inter-population exchange
 model, in a cluster of 16 nodes, each one with 16 Intel(R) Xeon(R) CPU E5520
@2.27GHz processors, 16 GB RAM, Broadcom NetXtreme II BCM5716 1000Base-T (C0) PCI Express network cards, CentOS 6.8 and Java Version 1.8.0\_80. Islands have been first distributed among the nodes. That means that, for example, the run of 8 islands requires 8 different nodes, instead of the usage of the 8 processors of a single node.

@@

%EXPLANATION OF THE RESULTS
This can be explained comparing the number of non-dominated solutions
in each island and the average number of generations (Table
\ref{tab:sols512} and Table \ref{tab:sols2048}). With both chromosome
lengths, increasing the number of islands implies more generations
with all the methods (logically, as there are less individuals in each
island). But also, the SLA methods
approach to the number of generations with the baseline when
increasing the number of islands. However, with $L=2048$ the migration
time is big enough  to not improve the number of generations
with respect to the baseline, even in the lower number of islands, as
it happens with $L=512$. Therefore, more generations do not
necessarily improve the solution of the global PF, but focusing on
different elements of the chromosome. As indicated previously, every
island is unaware of the other islands PFs, and they are trying to
optimize their solutions independently. With respect to the average
number of solutions in each island, there is a clear difference 
 with the baseline with $L=512$, where this value is in the most
of the cases, less than a half. The number of non-dominated solutions
also implies a better Spread indicator, where the baseline 
 obtains better (or not significantly different) values in the majority of the compared configurations. %TENGO QUE COMENTAR MAS LOS RESULTADOS Y
                              %ENTENDERLO TODO Y JUSTIFICARLO
                              %FINAMENTE
%Almost always no es muy preciso - JJ Pablo: arreglao

@@

\section*{Acknowledgements}
This work has been partially funded by projects TIN2014-56494-C4-3-P and TIN2015-67020-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR),  and  MSTR (PRY142/14, Fundaci\'on P\'ublica Andaluza Centro de Estudios Andaluces en la IX Convocatoria de Proyectos de Investigaci\'on).

>>>>

Specifically, in this paper, a new method to
automatically adapt the size of the overlapping, or shared sections,
of the individuals has been compared to other collaborative
evolutionary techniques. Different number of islands, chromosome sizes
and problems have been used. The analysis of the obtained experimental
results, by using different metrics, shows that our approach can provide statistically significant improvements with
respect to the base algorithm.%, being the method to automatically adapt
%the overlapping size the one to obtain the best performance during the
%same execution time. 
 Results also show that the relation between the number
of islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables. 
\end{abstract}

@@

The rest of the paper is organized as follows: after the State of the Art in distributed and co-evolutionary MOEAs, 
the used methodology and the compared algorithms are described in Section \ref{sec:methodology}. 
Then, the results of the experiments are presented (Section \ref{sec:results}), Finally, the conclusions and future work are discussed.

@@

Since the early 2000s distributed and parallel EAs
 have been used mostly to leverage systems such
as clusters or grids \citep{Talbi08Parallel}, since a first-order
approach to EA parallelization is as easy as dividing the population
and add a simple communication mechanism among them. But in the case of
MOEAs, the distribution and parallelization is not as direct as in
single-objective EAs. This is because in different steps of the
algorithm a whole set of dependent solutions, the Pareto Front, should
be managed as a whole, spending time in gathering all individuals from
the different processors or islands.

To solve this issue, some authors have proposed using Master-Slave approaches. For example,  \citep{Durillo08masterslave} compared different master-slave approaches: synchronous generational, asynchronous generational and asynchronous steady-state, being the latter the most promising option. On the other hand, the method proposed by Hiroyasu \citep{Hiroyasu07discussion} generates offspring depending on the computational power.

%This is an improvement over the previous? What is the relationship? FERGU: linked
Other kind of approaches have also been explored. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. In that paper, authors concluded that dividing the search space is a good idea, although achieving this is not trivial. The division of the search space has been explored by other researchers, for example, dividing the population in elite and search sub-populations \citep{Wang09parallel}, or separating it in processors by objective \citep{Xiao03specialized}. Other authors, such as Martens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

@@

\section{Methodology} % This should be called
                                % methodology. The object of the
                                % section is to explain the
                                % methodology used to test the
                                % problem, in order to do that we have
                                % tested different versions that prove
                                % that our methodology works for the
                                % precise reasons we think it works- JJ FERGU: Done, and explained
\label{sec:methodology}

@@

\subsection{Proposed methods}
In this paper we analyse several SLA algorithms we have
implemented by using a parallel multi-objective baseline algorithm
as a basis for comparisons. Our proposed methods, using different overlapping schemes, %methods or method?
                                %Which methods? - JJ Fergu: clarifying
are based on NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. % This is SoA stuff - JJ FERGU: Yes, I added it here againg because a reviewer wasn't very attentive
Therefore, we have used a basic distributed NSGA-II algorithm without
overlapping sections as the baseline (B).

%This probably should deserve a "Algorithm" environment too - JJ FERGU: You mean the LaTeX one? I've used it in the pseudocode figures
This basic distributed algorithm spreads the population among $P$
islands; after a fixed number of generations,
one individual in a given island is migrated to another random island,
thus avoiding to synchronize the global Pareto Front (PF) every
certain number of generations as other methods described in Section 2
do. At the end of the run, the Pareto Fronts of all islands are
aggregated into a new one, and the quality measures are evaluated. 
Different  SLA alternatives can be devised to evolve the
subpopulations according to the decision space to be explored by each
island. Here we will consider three methods in which each island only
performs crossover and mutation in specific sections of the
individuals.

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier (from 1 to $P$). Note that, differently from other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel}, all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with decomposable and no decomposable problems. %Don't understand what you say
                                %here. What you are saying is that our
                                %approach can't deal with decomposable
                                %problems, actually. - JJ FERGU: you are right, fixed

@@

\subsubsection{SLA with disjoint islands (D)} 
In this approach, each individual of size $L$ is split into $P$ chunks
of size $L/P$. Every island $p$ only performs crossover and mutation
on the $p_{th}$ part of the individuals. Figure \ref{fig:disjoint}
describes this approach.   
%
\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandDisjoint.jpg}
\caption{SLA with disjoint islands (D): every island $p$ only modifies the $p_{th}$ components (in grey) of the individuals.}
\label{fig:disjoint}
\end{figure*}

\subsubsection{SLA with overlapping islands (O)}
This approach is similar to the previous one, but every island also uses the $p+1$ and $p-1$ (module size) chunks of the individual for crossover and migration. Therefore, some kind of overlapping of the crossed and mutated parts exists between islands. Figure \ref{fig:overlapping} shows the affected parts of the individuals in
each island.

@@

\subsubsection{Adaptive SLA with overlapping islands (A)} 
As in the previous method, the sections to address by the operators are overlapped, but
instead of using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.
% That's not adaptive. It's simply variable - JJ FERGU: it depends on the number of islands. What term do you suggest?
As a first
approach to automatically calculate this value, the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands were $>$8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to create the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. As mentioned, this is only a first approach to obtain $c$, and future work will study new methods to calculate it. Figure \ref{fig:adaptive} explains this method.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%  EXPERIMENTS AND RESULTS  %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Experimental setup}

This subsection describes the quality indicators used and the experimental setup. The chosen quality indicators, are:

@@

The ECJ framework \citep{ECJ} has been used to run the
experiments. Specific operators have been developed as new modules for
ECJ, and they can be downloaded from our GitHub repository under a
LGPL V3 License
\footnote{\url{https://github.com/hpmoon/hpmoon-islands}}. 
%\footnote{\url{https://ANONYMOUS-REPOSITORY}}. 
The
island model has been executed asynchronously, using the ECJ distributed inter-population exchange
 model, in a cluster of 16 nodes, each one with 16 Intel(R) Xeon(R) CPU E5520
@2.27GHz processors, 16 GB RAM, Broadcom NetXtreme II BCM5716 1000Base-T (C0) PCI Express network cards, CentOS 6.8 and Java Version 1.8.0\_80. Islands have been first distributed among the nodes. That means that, for example, running with 8 islands requires 8 different nodes, instead of the usage of the 8 processors of a single node.

\section{Results}
\label{sec:results}

@@

%EXPLANATION OF THE RESULTS
This can be explained comparing the number of non-dominated solutions
in each island and the average number of generations (Table
\ref{tab:sols512} and Table \ref{tab:sols2048}). With both chromosome
lengths, increasing the number of islands implies more generations
with all the methods (logically, as there are less individuals in each
island). But also, the SLA methods
approach to the number of generations with the baseline when
increasing the number of islands. However, with $L=2048$ the migration
time is big enough  to not improve the number of generations
with respect to the baseline, even in the lower number of islands, as
it happens with $L=512$. Therefore, more generations do not
necessarily improve the solution of the global PF, but focusing on
different elements of the chromosome. As  previously stated, every
island is unaware of the other islands PFs, and they are trying to
optimize their solutions independently. With respect to the average
number of solutions in each island, there is a clear difference 
 with the baseline with $L=512$, where this value is in the most
of the cases, less than a half. The number of non-dominated solutions
also implies a better Spread indicator, where the baseline 
 obtains better (or not significantly different) values in the majority of the compared configurations. %TENGO QUE COMENTAR MAS LOS RESULTADOS Y
                              %ENTENDERLO TODO Y JUSTIFICARLO
                              %FINAMENTE
%Almost always no es muy preciso - JJ Pablo: arreglao

@@

\section*{Acknowledgements}
This work has been partially funded by projects TIN2014-56494-C4-3-P and TIN2015-67020-P (Spanish Ministry of Economy and Competitivity).
